Redshift Evolution of the Type Ia Supernova Population Stretch Distribution

### 2020-02-03 Nicolas.tex
New plot

\author{N. Nicolas
\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
\inst{1}
    \and M. Rigault
\thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
\inst{2}
    \and R. Graziani\inst{2}
    \and M. Briday\inst{1}
    \and Y. Copin\inst{1}
    \and Y. Kim\inst{1}
}

@@

For the last decades, numerous analysis have studied the relation between SNe~Ia
and host properties\nn{,} finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar-mass, SNe~Ia from high-mass host
\nn{being brighter on average}
\cite[e.g.][]{kelly2010,sullivan2010,childress2013,betoule2014,rigault2018}.
This mass-step correction is currently used in cosmological analyses
\citep[e.g.][]{betoule2014,scolnic2018a}, including for deriving H$_0$
\citep{riess2016,riess2019}. Yet, the underlying connection between the SN and
the host remains unclear when using global properties such as the host stellar
mass, which raises the question of the accuracy of the correction. More
recently, studies have used the local SN environment to probe more direct
connections between the SN and its environments \citep{rigault2013}\nn{,}
showing that local age-tracers such as the local specific star formation rate or
the local color are more strongly correlated with the standardized SN magnitude
\citep{rigault2018,roman2018}, suggesting the age to be the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology because the redshift drift magnitude correction to apply might
strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
importance of local SN environmental study remains highly debated
\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
derivation of H$_0$ \cite{jones2015,riess2016,riess2018,rose2019}.

@@

The concept of the SNe~Ia age dichotomy arised when the SN~Ia rate has first
been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
have shown that the relative SNe~Ia rate in galaxies could only be explained if
two populations existed, one young, following the host star formation activity,
and one old following the host stellar mass (the so called “prompt and delayed“
or “A+B“ model). In \cite{rigault2018} we used the specific star formation rate
at the SN location (Local sSFR or LsSFR) to classify which are the prompts
(those with a high LsSFR) and which are the delayed (those with low LsSFR).
Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
to be strongly correlated with the SN host properties \citep{hamuy1996,
hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014}. Following the “A+B“ model and the connection between SN
stretch and host properties, \cite{howell2007} first discussed the potential
redshift drift of the SN stretch distribution. In this paper we revisit this
analysis with the most recent SNe~Ia dataset..

@@

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}.  A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SN~Ia that a
magnitude-limited survey will miss are the reddest and the lowest-stretch ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely.

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
        The individual SN data are shown as blue dots and a 2D histogram is
        shown in gray to highlight point density. The ellipse
        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \citep{scolnic2018a}. \nn{The lines represent
        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
    for $z=0.60$ and $z=0.70$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
    above these lines are those for which $m > m_{lim}$, meaning they can't be
    observed.}}
    \label{fig:maglim}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{Article_figures/hist_surveys_cuts.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the pantheon dataset, respectively. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
full(light) color responds to the conservative(nominal) cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). \nn{Their spectroscopic follow-up were only done on a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
SN population. The SNf data is particularly useful for studying SN property
drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
from Pantheon follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{FIND BACK THE REF}, see
table~\ref{tab:sample}.

@@

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
        the local specific star formation rate (LsSFR) for SNf data used in this
        analysis. The color corresponds to the probability for the SNe~Ia to be
        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p(y)$; see
        \cite{rigault2018}. \textit{Right}: histogram of $p(y)$-weighted
        marginalization of the SN stretch. On both panel, we illustrate the
        Base model fitted on the SNf data. The young and old population
        contributions are shown in purple and yellow, respectively.}
    \label{fig:stretchlssfr}
\end{figure*}

@@

\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
    \mathrm{d}x^{i}_{1}, \nn{y^i}}\right),
\end{equation}

\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
    $y^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
    respectively. Depending on the data available, there are two ways to fit
these 5 free parameters}.

@@

\section{\nn{BBC discussion}}
\label{sec:bbc}
\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
    distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
    modelisation. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the best of all
    the non-drift model, but still $\Delta \mathcal{L}=-11.5$ worse than our
    best model, it is penalized by requiring 15 free parameters, 3 per sample.
    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
    highlighted that, because current surveys span limited redshift ranges, the
    per sample approach somewhat accounts for unmodeled redshift drifts. We
    stress here that, as measurements of modern surveys cover larger redshift
    ranges to reduce calibration systematic  uncertainties, this is becoming
    less true, notably for PS1, DES and, soon, LSST. Already with the current
    surveys, the BBC sample-based technique is excluded to be an as good
    representation of the data as our best model ($p=4\times 10^{-8}$). A more
    detailed discussion of the consequence of this results for cosmology follows
in section~\ref{sec:bbc}}

\mr{We report in table~\ref{tab:bbc} the samples' $\sigma^-$, $\sigma^+$ and
    $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{the} somewhat
    simple construction of our selection-bias free sample, see
    section~\ref{sec:sample}. If we were to use the \cite{scolnic2016} and
    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
    modeling would go from -32 to -47. \nn{It would seem that our minimization
    of these parameters is closer to our Base model than theirs}. Furthermore,
    fixing the Base model parameters to those derived using SNf data only (top
    row table.~\ref{tab:modelresults}) and using the BBC parameters from the
    aforementioned literature for PS1, SDSS and SNLS, we find
    $\Delta\mathrm{AIC}=-31.7$. Removing the SNf data reduces this difference to
    $\Delta\mathrm{AIC}=-5$. Such a reduction is expected for two reasons.
    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
    redshift level arm to probe the impact of potential redshift drifts.
    Second, the SNf data are better at discriminating the individual stretch
    distributions by age since the probability for a SN to be young $p_y{}^i$ is
    known while other samples rely on $\delta(z^i)$ which only is the fraction
    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for
    $p_y{}^i$. While in any case the Base-model is favored, the relative weight
    of a SNf sample highlights the importance of adding non-targeted nearby
surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
graham2019} to unambiguously confirm our result.}

@@

\mr{We illustrate Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
underlying stretch distribution between the BBC modeling and our Base drifting
model for the PS1 sample. By definition, our model is bimodal and the relative
amplitude of each mode depends on the fraction of old SNe~Ia in the sample
--~the highest the fraction of old SNe~Ia, the highest the amplitude of the
low-stretch mode~-- and is therefore redshift dependent. As expected the two
approaches strongly differ in modeling the negative part of the SN stretch
distribution. The BBC asymmetric distribution goes in the middle of the bimodal
distribution, largely over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
largely under-estimating it at $x_1\sim-1.7$ for typical PS1 SN redshifts. This
means that the bias-correction of the individual SN stretch
$\overline{\delta}_{x_1}$ \citep{kessler2017} and consequently the magnitude
bias correction $\mu_B$ is most likely inaccurate. However, given the complexity
of the BBC analysis, a full study using our Base model given
eq.~\ref{eq:stretchz} in place of the sample-based asymmetric modeling would be
necessary to have an exact understanding of the effect this inaccurate modeling
has on cosmology.}

@@

Following observations made in \cite{rigault2018} we introduced a redshift drift
modeling that depends on the expected fraction of young and old SNe~Ia as a
function of redshift, each age-population having its own underlying stretch
distribution. We studied various suites of distribution as well as non-drifting
modeling. We also closely studied the stretch distribution modeling made in
current version of the BCC algorithm used for Malmquist bias correction by all
recent Dark energy or Hubble constant cosmological analyses.

@@

    \item Non-drifting models are excluded to be as good descriptions of the
        data \nn{as} our Base modeling. This model assumes that: (1) the young
        population has an unimodal Gaussian stretch distribution while the old
        population stretch one is bimodal, one mode being the young one; (2) the
        evolution of the relative fraction of young and old SNe~Ia follows the
        prediction made in \cite{rigault2018}.

    \item Given 1. we conclude that the SNe~Ia is indeed drifting, as already
        suggested by \citep[e.g.][]{howell2007}.

    \item The BBC approach, which assumes each survey to have its own asymmetric
        Gaussian distribution, is excluded to be an as good description of the
        data as our drifting model ($p=4\times10^{-8}$). Hence, the BBC
        sample-based approach does not accurately account for redshift drift as
        suggested by \cite{scolnic2016} and \cite{scolnic2018a}.

    \item An inaccurate underlying stretch distribution modeling is estimated to
        bias the mean standardized SNe~Ia on the order of a few percent, which
        is comparable in scale to the observational signature of exotic forms of
        dark energy.

    \item We suggest to use the following stretch population modeling:
    \begin{align}
    \label{eqconclusion:stretchz}
        \Delta\left(x_1\,|\,z \right) =
        &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1)\,+\nonumber\\
        (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1) +
        (1-a)\times\mathcal{N}(\mu_2,\sigma_2)\right]\nonumber
    \end{align}
    with: $a=0.50$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
    $\sigma_2=0.56$, see table~\ref{tab:modelresults} and using the
    age-population drift modelization $\delta(z)$ defined in \cite{rigault2018}
    with $K=0.87$:
    \begin{align}
        \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\nonumber
    \end{align}
\end{enumerate}

@@

@@

>>>>

\author{N. Nicolas
\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
\inst{1}
    \and M. Rigault
\thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
\inst{2}
    \and R. Graziani\inst{2}
    \and M. Briday\inst{1}
    \and Y. Copin\inst{1}
    \and Y.-L. Kim\inst{1}
}

@@

For the last decades, numerous analysis have studied the relation between SNe~Ia
and host properties\nn{,} finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar-mass, SNe~Ia from high-mass host
\nn{being brighter on average}
\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2018, kim19}.
This mass-step correction is currently used in cosmological analyses
\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving H$_0$
\citep{riess2016, riess2019}. Yet, the underlying connection between the SN and
the host remains unclear when using global properties such as the host stellar
mass, which raises the question of the accuracy of the correction. More
recently, studies have used the local SN environment to probe more direct
connections between the SN and its environments \citep{rigault2013}\nn{,}
showing that local age-tracers such as the local specific star formation rate or
the local color are more strongly correlated with the standardized SN magnitude
\citep{rigault2018, roman2018, kim18}, suggesting the age to be the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology because the redshift drift magnitude correction to apply might
strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
importance of local SN environmental study remains highly debated
\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.

@@

The concept of the SNe~Ia age dichotomy arised when the SN~Ia rate has first
been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
have shown that the relative SNe~Ia rate in galaxies could only be explained if
two populations existed, one young, following the host star formation activity,
and one old following the host stellar mass (the so called “prompt and delayed“
or “A+B“ model). In \cite{rigault2018} we used the specific star formation rate
at the SN location (Local sSFR or LsSFR) to classify which are the prompts
(those with a high LsSFR) and which are the delayed (those with low LsSFR).
Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
to be strongly correlated with the SN host properties \citep{hamuy1996,
hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014, kim19}. Following the “A+B“ model and the connection between SN
stretch and host properties, \cite{howell2007} first discussed the potential
redshift drift of the SN stretch distribution. In this paper we revisit this
analysis with the most recent SNe~Ia dataset..

@@

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SN~Ia that a
magnitude-limited survey will miss are the reddest and the lowest-stretch ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely \nn{\citep[e.g.,][]{kim19}}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
        The individual SN data are shown as blue dots and a 2D histogram is
        shown in gray to highlight point density. The ellipse
        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \cite{scolnic2018a}. \nn{The lines represent
        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
    above these lines are those for which $m > m_{lim}$, meaning they can't be
    observed.}}
    \label{fig:maglim}
\end{figure}

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the pantheon dataset. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
full(light) color responds to the conservative(nominal) cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). \nn{Their spectroscopic follow-up were only done on a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
SN population. The SNf data is particularly useful for studying SN property
drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
from Pantheon follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{strolger04}, see
table~\ref{tab:sample}. \nn{We present the stretch distribution, redshift histogram
and stretch cdf of these five surveys Fig.~\ref{fig:sample}}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75.pdf}
    \caption{\nn{\textit{Main}: Stretches of each survey in full and empty markers for the
    conservative and nominal cuts, respectively, as a function of redshift. \textit{Right}:
    normed CDF of each survey's stretch distribution. \textit{Top}: combined redshift 
    histogram of all the data, in dark gray for the conservative cuts;
    in light gray for the nominal cuts.}}
    \label{fig:sample}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
        the local specific star formation rate (LsSFR) for SNf data used in this
        analysis. The color corresponds to the probability for the SNe~Ia to be
        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p_y$; see
        \cite{rigault2018}. \textit{Right}: histogram of $p_y$-weighted
        marginalization of the SN stretch. On both panel, we illustrate the
        Base model fitted on the SNf data. The young and old population
        contributions are shown in purple and yellow, respectively.}
    \label{fig:stretchlssfr}
\end{figure*}

@@

\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
    \mathrm{d}x^{i}_{1}, \nn{y^i}}\right),
\end{equation}
\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
    $y^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
    respectively. Depending on the data available, there are two ways to fit
these 5 free parameters}.

@@

\section{\nn{BBC discussion}}
\label{sec:bbc}
\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
    distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
    modelisation. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the best of all
    the non-drift model, but still $\Delta \mathcal{L}=-11.5$ worse than our
    best model, it is penalized by requiring 15 free parameters, 3 per sample.
    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
    highlighted that, because current surveys span limited redshift ranges, the
    per sample approach somewhat accounts for unmodeled redshift drifts. We
    stress here that, as measurements of modern surveys cover larger redshift
    ranges to reduce calibration systematic  uncertainties, this is becoming
    less true, notably for PS1, DES and, soon, LSST. Already with the current
    surveys, the BBC sample-based technique is \nn{ruled out} to be an as good
    representation of the data as our best model ($p=4\times 10^{-8}$). A more
    detailed discussion of the consequence of this results for cosmology follows
in section~\ref{sec:bbc}}

\mr{We report in table~\ref{tab:bbc} the samples' $\sigma^-$, $\sigma^+$ and
    $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{our} construction
    of \nn{a} selection-bias free sample, see
    section~\ref{sec:sample}. If we were to use the \cite{scolnic2016} and
    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
    modeling would go from -32 to -47. \nn{It would seem that our minimization
    of these parameters is closer to our Base model than theirs}. Furthermore,
    fixing the Base model parameters to those derived using SNf data only (top
    row table.~\ref{tab:modelresults}) and using the BBC parameters from the
    aforementioned literature for PS1, SDSS and SNLS, we find
    $\Delta\mathrm{AIC}=-31.7$. Removing the SNf data reduces this difference to
    $\Delta\mathrm{AIC}=-5$. Such a reduction is expected for two reasons.
    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
    redshift \nn{lever} arm to probe the impact of redshift drifts.
    Second, the SNf data are better at discriminating the individual stretch
    distributions by age since the probability for a SN to be young $p_y{}^i$ is
    known while other samples rely on $\delta(z^i)$ which is the fraction
    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for
    $p_y{}^i$. While in any case the Base-model is favored, the relative weight
    of the SNf sample highlights the importance of adding non-targeted nearby
surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
graham2019} to unambiguously confirm our \nn{model}.}

@@

\mr{We illustrate Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
underlying stretch distribution between the BBC modeling and our Base drifting
model for the PS1 sample. Our model is bimodal and the relative
amplitude of each mode depends on the fraction of old SNe~Ia in the sample
--~the highest the fraction of old SNe~Ia, the highest the amplitude of the
low-stretch mode~-- and is therefore redshift dependent. As expected the two
approaches strongly differ in modeling the negative part of the SN stretch
distribution. The BBC asymmetric distribution goes in the middle of the bimodal
distribution, largely over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and
largely under-estimating it at $x_1\sim-1.7$ for typical PS1 SN redshifts. This
means that the bias-correction of the individual SN stretch
$\overline{\delta}_{x_1}$ \citep{kessler2017} and consequently the magnitude
bias correction $\mu_B$ is most likely inaccurate. However, given the complexity
of the BBC analysis, a full study using our Base model given
eq.~\ref{eq:stretchz} in place of the sample-based asymmetric modeling would be
necessary to have an exact understanding of the effect this inaccurate modeling
has on cosmology.}

@@

Following observations made in \cite{rigault2018} we introduced a redshift drift
modeling \nn{which} depends on the expected fraction of young and old SNe~Ia as a
function of redshift, each age-population having its own underlying stretch
distribution. We have studied various suites of distribution\nn{, including} non-drifting
modeling. We also studied the stretch distribution modeling \nn{of the}
current version of the BBC algorithm used for Malmquist bias correction by all
recent Dark energy or Hubble constant cosmological analyses.

@@

    \item Non-drifting models are excluded to be as good descriptions of the
        data \nn{as} our Base model. This model assumes that: (1) the young
        population has an unimodal Gaussian stretch distribution while the old
        population stretch one is bimodal, one mode being the young one; (2) the
        evolution of the relative fraction of young and old SNe~Ia follows the
        prediction made in \cite{rigault2018}.

    \item Given 1. we conclude that the SNe~Ia \nn{stretch} is indeed drifting, as
        already suggested by \citep[e.g.][]{howell2007}.

    \item The BBC approach, which assumes each survey to have its own asymmetric
        Gaussian distribution, is excluded to be an as good description of the
        data as our drifting model ($p=4\times10^{-8}$). Hence, the BBC
        sample-based approach does not accurately account for redshift drift \nn{contrary
        to the suggestion of} \cite{scolnic2016} and \cite{scolnic2018a}.

    \item An inaccurate underlying stretch distribution modeling is estimated to
        bias the mean standardized SNe~Ia \nn{to} the order of a few percents, which
        is comparable in scale to the observational signature of exotic forms of
        dark energy.

    \item We suggest to use the following stretch population modeling:
        \begin{align}
            \label{eqconclusion:stretchz}
            \nn{X_1\left(z \right) =
            } &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
            (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
            (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
        \end{align}
    with: $a=0.50$, $\mu_1=0.37$, $\mu_2=-1.22$, $\sigma_1=0.61$,
    $\sigma_2=0.56$, see table~\ref{tab:modelresults} and using the
    age-population drift modelization $\delta(z)$ defined in \cite{rigault2018}
    with $K=0.87$:
    \begin{align}
        \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\nonumber
    \end{align}
\end{enumerate}

@@

\bibitem[Kim et al.(2018)]{kim18} Kim, Y.-L., Smith, M., Sullivan, M., et al.\ 2018, \apj, 854, 24

\bibitem[Kim et al.(2019)]{kim19} Kim, Y.-L., Kang, Y., \& Lee, Y.-W.\ 2019, Journal of Korean Astronomical Society, 52, 181

@@

\bibitem[Strolger et al.(2004)]{strolger04} Strolger, L.-G., Riess, A.~G., Dahlen, T., et al.\ 2004, \apj, 613, 200
